---
title: "Interactive Simulation Monitoring"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive Simulation Monitoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Interactive provides a Shiny monitor for a simulation driven by a Runner. The
monitor is responsible for displaying data and status; the Runner remains
responsible for executing simulation and aggregate functions.

Before creating the monitor, define the plotting callbacks that describe what
should be displayed. `plot_fun` is required: it receives the accumulated output
from the current design condition and must return the live plot. `agg_plot_fun`
is optional: define it when you also want a second plot showing the aggregated
results accumulated across completed design conditions. It receives the
aggregated output and is used to populate the **Aggregated** tab.

The callbacks are called with these arguments:

```{r callback-signatures-intro, eval=FALSE}
plot_fun(data, step)
agg_plot_fun(agg_x, step)
```

Here `data` is the current-condition output, `agg_x` contains aggregated output
from all completed conditions, and `step` is the number of updates received.
The plotting callbacks should return a plot object that Shiny can render. The
Runner does not create these plots automatically; it calls the callbacks as the
simulation produces results.

## Define a simulation

This example uses three design conditions. Each replication returns a sample,
and the aggregate function computes its mean.

```{r setup, include=FALSE}
interactive_deps <- all(vapply(
  c("later", "shiny"),
  requireNamespace,
  logical(1),
  quietly = TRUE
))

knitr::opts_chunk$set(
  eval = interactive_deps,
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r define-runner}
runner <- Rsimcity::Runner$new("interactive-demo")
runner$params <- list(N = 80)
runner$design <- list(mu = c(-1, 0, 1))

runner$step <- function(N, mu) {
  data.frame(value = stats::rnorm(N, mean = mu))
}

runner$aggregate <- function(data, mu) {
  data.frame(mu = mu, mean_value = mean(data$value))
}
```

## Define the plots

The live plot receives the accumulated output for the current condition. The
aggregate plot receives the results from all completed conditions.

```{r define-plots}
live_plot <- function(data, step) {
  ggplot2::ggplot(data, ggplot2::aes(x = value)) +
    ggplot2::geom_histogram(bins = 25, fill = "#2C7FB8", colour = "white") +
    ggplot2::labs(
      title = sprintf("Current condition, update %d", step),
      x = "Simulated value",
      y = "Count"
    ) +
    ggplot2::theme_minimal()
}

aggregate_plot <- function(data, step) {
  ggplot2::ggplot(data, ggplot2::aes(x = mu, y = mean_value)) +
    ggplot2::geom_hline(yintercept = 0, colour = "grey70") +
    ggplot2::geom_line(colour = "#D95F0E") +
    ggplot2::geom_point(size = 2, colour = "#D95F0E") +
    ggplot2::labs(
      title = "Completed conditions",
      subtitle = sprintf("Latest update: %d", step),
      x = "Population mean",
      y = "Observed mean"
    ) +
    ggplot2::theme_minimal()
}
```

## Connect the monitor

Passing agg_plot_fun creates a second Aggregated tab. The start callback
connects the monitor to Runner$experiment_interactive().

```{r create-monitor}
monitor <- Rsimcity::Interactive$new(
  plot_fun = live_plot,
  agg_plot_fun = aggregate_plot,
  title = "Interactive simulation demo",
  plot_refresh_ms = 250,
  plot_height = "600px",
  max_rows = 2000
)

monitor$set_start_fun(function(obj) {
  runner$experiment_interactive(
    interactive = obj,
    Rep = 20,
    delay = 0.05,
    reset = FALSE
  )
})
```

## Run the app

The following command opens the monitor in the RStudio Viewer pane. It is
marked eval=FALSE because launching a Shiny app would block vignette
rendering. Run it interactively in the Console after evaluating the preceding
chunks.

```{r run-monitor, eval=FALSE}
monitor$run(viewer = "pane")
```

The app initially shows a waiting plot. Press **Start** to begin the
simulation. The controls have these effects:

- **Start** resets the monitor and invokes start_fun.
- **Stop** pauses future simulation updates.
- **Continue** resumes a paused simulation.
- **Reset** clears live and aggregated data.

During execution, Runner$experiment_interactive() calls monitor$update() after
each replication. When a design condition finishes, it calls
monitor$update_agg(); this is what updates the Aggregated tab. The two callbacks
are called as follows:

```{r callback-signatures, eval=FALSE}
live_plot(data, step)
aggregate_plot(agg_x, step)
```

## Live-only monitoring

An aggregated plot is optional. If only the current condition is relevant,
omit agg_plot_fun:

```{r live-only, eval=FALSE}
live_monitor <- Rsimcity::Interactive$new(plot_fun = live_plot)
live_monitor$set_start_fun(function(obj) {
  runner$experiment_interactive(interactive = obj, Rep = 20)
})
live_monitor$run(viewer = "pane")
```

The same monitor can also be opened in the system browser with
viewer = "browser", or created without opening a viewer with viewer = "none".
