---
title: "Introduction to Rsimcity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Rsimcity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This short vignette demonstrates a minimal, non-interactive use of the
Rsimcity Runner to generate simple simulation results and aggregate them.

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(Rsimcity)
```

## Minimal example

Create a runner, define parameters and design, add a simulation step, then
run the experiment with a small number of replications:

```{r example}
runner <- Rsimcity::Runner$new("example")
runner$params <- list(N = 100)
runner$design <- list(mu = c(0, 1, 2))

runner$step <- function(N, mu) {
  data.frame(x = stats::rnorm(N, mean = mu))
}

# Aggregate function: compute mean of x for the condition
runner$aggregate <- function(data, mu) {
  data.frame(mu = mu, mean_x = mean(data$x))
}

# Run the experiment (Rep = number of replications per condition)
res <- runner$experiment(Rep = 10, progress = FALSE)

head(res)
```

You can edit and extend this vignette to include Interactive usage or plots.
