---
title: "Introduction to Rsimcity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Rsimcity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This short vignette demonstrates a minimal, non-interactive use of the
Rsimcity Runner to generate simple simulation results and aggregate them.

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
library(Rsimcity)
```

## Minimal example

Create a runner, define parameters and design, add a simulation step, then
run the experiment with a small number of replications:

```{r example}
runner <- Rsimcity::Runner$new("example")
runner$params <- list(N = 100)
runner$design <- list(mu = c(0, 1, 2))

runner$step <- function(N, mu) {
  data.frame(x = stats::rnorm(N, mean = mu))
}

# Aggregate function: compute mean of x for the condition
runner$aggregate <- function(data, mu) {
  data.frame(mu = mu, mean_x = mean(data$x))
}

# Run the experiment (Rep = number of replications per condition)
res <- runner$experiment(Rep = 10, progress = FALSE)

head(res)

```

## Plotting split simulation results

plot_by_splits() accepts several response columns and can use one grouping
variable for color and linetype, or two grouping variables for color and
linetype separately. The first two split variables form the columns and rows
of the plot; additional split variables return a named list of plots.

~~~{r plot-by-splits}
plot_data <- expand.grid(
  x = c(1, 2, 3),
  model = c("A", "B", "C"),
  scenario = c("baseline", "alternative"),
  method = c("actual", "adjusted"),
  rep = 1:2,
  stringsAsFactors = FALSE
)

plot_data$estimate <- with(
  plot_data,
  x / 3 +
    ifelse(model == "B", 0.05, ifelse(model == "C", 0.10, 0)) +
    ifelse(scenario == "alternative", 0.03, 0) +
    ifelse(method == "adjusted", 0.02, 0) +
    rep / 100
)
plot_data$target <- with(plot_data, x / 3 + rep / 100)

figure <- plot_by_splits(
  data = plot_data,
  xvar = "x",
  yvar = c("estimate", "target"),
  zvar = "method",
  splits = c("model", "scenario"),
  title = c("Baseline", "Alternative"),
  titles = "top",
  xlabel = "Simulation size",
  ylabel = "Value",
  color_labels = c(estimate = "Estimate", target = "Target"),
  linetype_labels = c(actual = "Actual", adjusted = "Adjusted")
)

figure
~~~

For top row titles, plot.title controls the title text and panel.spacing.y
controls the space around the title rows. These are ordinary ggplot2 theme
settings and can be changed after the plot is created:

~~~{r plot-by-splits-theme}
figure +
  ggplot2::theme(
    plot.title = ggplot2::element_text(size = 12),
    panel.spacing.y = grid::unit(10, "pt")
  )
~~~
