---
title: "Parallel computing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallel computing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
library(Rsimcity)
```

`Runner` supports two independent levels of parallelism, and it's important
not to mix them up:

- **Across design cells**: `runner$parallel` runs each condition (each row of
  the design grid) as its own background process, via the
  [future](https://future.futureverse.org) package.
- **Within a single step**: your own `step` function can parallelize its own
  work internally (e.g. with `future.apply::future_lapply()`), regardless of
  whether `runner$parallel` is `TRUE` or `FALSE`.

This vignette covers both, with minimal examples.

## Parallel across design cells: `runner$parallel`

By default, `runner$parallel` is `FALSE`, and `Runner$experiment()` runs each
design cell one after another, in the current R session:

`runner$step` isn't limited to an inline function -- it just needs to hold a
function, so any ordinary function you've already defined works just as well:

```{r sequential}
my_step <- function(N, mu) {
  data.frame(x = rnorm(N, mean = mu))
}

runner <- Rsimcity::Runner$new("sequential example")
runner$params <- list(N = 200)
runner$design <- list(mu = c(0, 1, 2))
runner$step <- my_step
runner$aggregate <- function(data, mu) data.frame(mu = mu, mean_x = mean(data$x))

res <- runner$experiment(Rep = 5, progress = FALSE)
res
```

Setting `runner$parallel <- TRUE` runs each design cell (each value of `mu`
here) in its own background process instead, using `future::multisession`
(the default, portable on Windows/macOS/Linux) or `future::multicore`
(Linux/macOS only, lighter-weight forked processes) via `runner$par_method`:

```{r parallel}
runner$parallel <- TRUE
runner$par_method <- "multisession"   # or "multicore" on Linux/macOS

res <- runner$experiment(Rep = 5, progress = FALSE)
res
```

Same `step`/`aggregate`/`design`, just flip `runner$parallel` on -- nothing
else about the code changes. This is worth doing when you have many design
cells and each one takes a noticeable amount of time on its own, since each
cell becomes one unit of parallel work; with only a handful of cheap cells
there's little to gain.

## Parallel within a step

Sometimes a single design cell does most of the work internally -- for
example, a `step` that searches for something (a required sample size, a
calibration constant...) by running many internal trials itself, rather than
relying on `Runner`'s own `Rep` for replication. In that case, parallelizing
*inside* the step is more useful than parallelizing across the (possibly few)
design cells.

This works together with `runner$parallel <- FALSE` (the default): set your
own `future::plan()` once, and call `future.apply::future_lapply()` (or any
other future-based tool) *inside* your `step` function. `Runner` leaves your
plan alone whenever `runner$parallel` is `FALSE`, so this is safe.

```{r nested}
future::plan(future::multisession)

## A step that internally runs 200 trials in parallel and returns their
## average, instead of relying on Runner's own Rep for replication.
one_step <- function(mu) {
  trials <- future.apply::future_lapply(1:200, function(i) {
    mean(rnorm(50, mean = mu))
  }, future.seed = TRUE)
  data.frame(mu = mu, avg = mean(unlist(trials)))
}

runner2 <- Rsimcity::Runner$new("nested parallel example")
runner2$parallel <- FALSE          # Runner itself stays sequential across cells...
runner2$design <- list(mu = c(0, 1, 2))
runner2$step <- one_step
runner2$aggregate <- function(data, mu) data.frame(mu = mu, avg = data$avg)

res2 <- runner2$experiment(Rep = 1, progress = FALSE)   # ...one_step does its own parallel work
res2

future::plan(future::sequential)   # good practice: shut the workers down when done
```

A few things worth knowing about combining the two levels:

- **Don't parallelize both levels at once.** If `runner$parallel <- TRUE`
  *and* your `step` also spawns its own parallel workers, you get parallelism
  nested inside parallelism -- each of `Runner`'s background processes would
  try to launch its own set of workers, oversubscribing your CPU cores. Pick
  one level: either `runner$parallel <- TRUE` with a plain (non-parallel)
  `step`, or `runner$parallel <- FALSE` with a `step` that parallelizes
  internally, as above.
- **`runner$parallel <- FALSE` never touches your `future::plan()`.**
  `Runner`'s own per-cell loop doesn't need a plan in that case (it just calls
  your `step` directly, once per cell), so whatever plan you set beforehand
  for your own nested calls is left exactly as you set it.
- Remember to call `future::plan(future::sequential)` (or otherwise close the
  workers) when you're done with a nested-parallel run, to free up the
  background processes it started.
